### 2015-09-30 draft1leslieandkenny.Rnw
Made some progress on model and graphics.

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\topmargin}{-1.5cm}
\setlength{\maxwidth}{6.5in}

\title{\vspace{-2.0cm}The effect of nitrogen application time on infection and development of the Wheat Streak Mosaic Virus (WSMV)}
\author{Leslie Gains-Germain, Kenny Flagg}
\date{Spring 2015}

\begin{document}
<<setup, include=FALSE, cache=FALSE,echo=FALSE>>=
require(knitr)
opts_chunk$set(fig.width = 10, fig.height = 4, out.width = '\\linewidth',
               out.height = '0.4\\linewidth', dev = 'pdf',
               concordance = TRUE,
               tidy.opts = list(width.cutoff = 60, keep.blank.line=FALSE),
               size = 'footnotesize')
options(replace.assign = TRUE, width = 112, digits = 3, max.print = '72',
        show.signif.stars = FALSE)

require(xtable)
#require(foreign)
#require(arm)
require(tabplot)
require(car)
require(lme4)
require(lattice)
@

<<data,echo=FALSE>>=
# Read and set up the data
fert2 <- read.csv('fert2.csv')

# Correctly order nitrogen application levels
fert2$n.trt <- relevel(fert2$n.application, 'FALL')

# If status=INOC is the most interesting case, make INOC the baseline
fert2$status2 <- relevel(fert2$status, 'INOC')

# Create a row factor
fert2$row <- with(fert2, interaction(variety, block))

# Combine nitrogen and inoc/cntl (for the heatmap labels)
fert2$n.status <- with(fert2, interaction(n.trt, status))
@

<<fakedata,echo=FALSE>>=
# Create some fake data with no differences anywhere
set.seed(983724)
fert.sim1 <- data.frame(
  'block' = c(rep('A', 30), rep('B', 30), rep('C', 30), rep('D', 30), rep('E', 30), rep('F', 30)),
  'variety' = c(rep('Var1', 6), rep('Var2', 6), rep('Var3', 6), rep('Var4', 6), rep('Var5', 6)),
  'n.application' = rep(c('FALL', 'FALL', 'EARLY SPRING', 'EARLY SPRING', 'LATE SPRING', 'LATE SPRING'), 30),
  'status' = rep(c('CNTL', 'INOC'), 90),
  'total' = rep(30, 180),
  'infected' = rnbinom(180, 4, 7/9))
#  'infected' = sample(fert2$infected, 180, replace = TRUE))
@

<<fakepre,echo=FALSE>>=
# Create a row factor
fert.sim1$row <- with(fert.sim1, interaction(variety, block))

# Correctly order nitrogen application levels, the default is alphabetical
fert.sim1$n.trt <- relevel(fert.sim1$n.application, 'FALL')

# If status=INOC is the most interesting case, make INOC the baseline
fert.sim1$status2 <- relevel(fert.sim1$status, 'INOC')
@

\maketitle
\tableofcontents

\newpage

\section{Introduction}
\noindent

\section{Statistical Experience and Assistance Needed}
\noindent Nar B. Ranabhat is a PhD candidate in the Department of Land Resources and Environmental Science. He is currently working on several projects regarding the spread and development of the Wheat Streak Mosaic Virus (WSMV) in Winter Wheat. Nar has taken Statistics $511$ and $512$ here at Montana State University, and he is currently auditing Statistics $448$. Nar has asked for our help in building a model, fitting it in R, and interpreting the results.

\section{Objectives and Questions}
\noindent Nar is interested in the effect of variety, nitrogen application time, and inoculation on the probability of virus infection in Winter Wheat. He is primarily interested in the effect of nitrogen application time, and he wants to assess evidence for all possible two and three way interactions. Nar's goal is to select the simplest possible model, and then use this model to describe the effects of the above variables. \\

We address the following questions in this report:
\begin{itemize}
\item Is there evidence that the interaction between nitrogen application and variety differs between inoculated and control plots?
\item Is there evidence that the nitrogen application effect varies across varieties? Is there evidence that the nitrogen application effect differs between inoculated and control?
\item Is there evidence that the variety effect differs between inoculated and control?
\item Conditional on the results to the above questions, how do we appropriately describe the estimated effects for variety, nitrogen application time, and inoculation on the probability of virus infection?
\end{itemize}

\section{Study Design/Data collection}

Nar conducted this experiment in a MSU field at the base of the Bridger Mountains. He chose one area in the field and divided it into six $31.5$ by $27.5$ meter blocks. The blocks were then divided into five rows, with $5$ meters of space between each row. Each row was randomly assigned to a variety of Winter Wheat with separate randomizations in each block. The rows then were divided into six $1.5$ by $5$ meter plots. Each plot was randomly assigned to one of six combinations of nitrogen application time and inoculation status, with separate randomizations in each row. The combinations were fall/inoculated, fall/control, early spring/inoculated, early spring/control, late spring/inoculated, and late spring/control. All plots were planted in the {\bf FALL?} and allowed to grow over the winter. The following {\bf September?}, $30$ plants were chosen from each plot to be taken to the lab and tested for presence of the virus. The selection of the plants from each plot was not random nor technically systematic. The {\bf undergraduate?} technicians were instructed to collect a sample of plants spread evenly throughout the field. {\bf different undergraduates doing picking? picking done all at the same time? is there visual evidence of the disease?} A total of $180$ plants were collected. The plants were then sent to the lab and screened for the virus via the ELISA procedure. This is a three year study. The data are collected for the first year, and the second year plants are waiting to be analyzed in the lab {\bf describe how they are stored during waiting time}. Planting for the third year will begin this fall {\bf I think?}.\\

\noindent Plots that were assigned to the inoculated treatment had five infected plants transplanted to the middle of the plot. These plants were infected {\bf as adults?} in the greenhouse by clipping an infected leaf to the healthy plant. The Wheat Streak Mosaic Virus is transmitted by the wheat curl mite, tiny organisms that are nearly invisible to the naked eye. They can easily move from leaf to leaf on a plant, and they are transported from plant to plant by the wind  (Sloderbeck $2008$). It is our understanding that once the infected mites were introduced to the study area, they were considered to be everywhere in the air and that all plants in the field were exposed to the infected mites.

\section{Recommendations}

\subsection{Binomial Generalized Linear Mixed Model}

\noindent Because the response variable is a Binomial count, and because
the design includes random assignment of variety to rows within blocks,
a binomial generalized linear mixed model is appropriate. This model
will estimate the probability of infection for each
\verb|variety:treatment:status| combination while controlling for
differences between blocks and between rows in the same block.

\noindent The model is
$$
y_i\sim\mathrm{Binomial}(30, p_i),
$$
$$
\mathrm{logit}(p_i)=\mathbf{x}_i\mathbf{\beta}+b_{0,j[i]}+b_{1,j[i],k[i]}
$$
where
\begin{itemize}
\item \(p_i\) is the probability of infection in the \(i\)th plot,
\item \(\mathrm{logit}(p_i)=\log\left(\dfrac{p_i}{1-p_i}\right)\) is the
natural logarithm of the odds of infection,
\item \(\mathbf{x_i}\) is a row vector containing indicator variables for
the variety, nitrogen application timing, and inoculation status of
observation \(i\),
\item \(\mathbf{\beta}\) is a column vector of fixed-effect coefficients, and
\item \(b_{0,j}\sim\mathrm{N}(0,\sigma^2_{b_0})\) and
\(b_{1,j,k}\sim\mathrm{N}(0,\sigma^2_{b_1})\) are random effects for block
and row, respectively.
\end{itemize}

\noindent The assumptions for this model are that plants within a plot
are independent and have the same probability of being infected, and
that plots within a row are endependent of each other.

\noindent Advantages of the Binomial GLMM include:
\begin{itemize}
\item It estimates and controls for variability between rows and blocks.
\item It avoids confounding variety effects with row effects and correctly
models the rows and nested within blocks.
\item Random effects can explain extra variation beyond what the
fixed-effect Binomial GLM can account for.
\end{itemize}

\noindent Disadvantages of this model include:
\begin{itemize}
\item It must be fit by numerically maximizing the likelihood with an
iterative algorithm. Including a large number of parameters (such as
third-order interaction coefficients) can cause the algorithm to fail
to converge.
\item The domain of the logit function is \((0, 1)\). If a large proportion
of the data are zeros, meaning an infection probability is near 0, the
algorithm will have difficulty estimating the coefficients. In this
situation, estimates may be unreliable or the algorithm may not converge.
\item We are not aware of any tools that can fit a Binomial GLMM with
overdispersion. If there are additional sources of variation that the
model does not account for, the standard error estimates will be too small.
\end{itemize}

\subsection{Graphical Data Exploration}

\noindent Before fitting a model, we recommend starting by creating plots
for exploratory data analysis. This is useful for seeing which interactions
may be present, and for identifying patterns that may represent violations
of assumptions or potential computational problems.

\noindent A good way to visually compare varieties, treatments, and
inoculation status is with boxplots of infection counts for each
combination of factor levels. The \verb|qplot| function in the
\verb|ggplot2| package is a simple way to do this.

\begin{figure}
\caption{Example interaction plot using \texttt{qplot}.}
<<interactionplot,echo=FALSE,message=FALSE>>=
require(ggplot2)
qplot(x = n.trt, y = infected, geom = 'boxplot', color = status,
      facets = .~variety, data = fert.sim1) +
  theme(axis.text.x = element_text(angle = 90))
@
\end{figure}

\noindent We noticed that infection counts of 0 were recored in 67 of the
180 plots in Nar's dataset. This is a large fraction of the data; if the
distribution of these zeros is related to the variables in the model then
it may be impossible to estimate coefficients for factor levels where
infection rates are nearly zero.

\noindent We recommend sorting the data and creating a heatmap to look
for patterns of zeros. Figure 2 shows an example using simulated data.
The most serious problems will come from the zeros, so we chose a contrasting
color (black) for plots with infection counts of 0.

\begin{figure}
\caption{Example heatmap visualization for identifying patterns of zeros.}
<<heatmap,echo=FALSE,message=FALSE>>=
## Heatmap with meaningful ordering

# Order the dataset, using the arrange() function from the dplyr package
require(dplyr)
fert.ordered <- arrange(fert.sim1, variety, block, status, n.trt)

# Create a matrix of the arranged responses
infected.arranged <- matrix(fert.ordered$infected, ncol = 6, byrow = TRUE)

# Set up two panels, right one for a legend
layout(t(1:2), widths = c(9, 1))

# Plot the heatmap, with zeros in black and segments separating the blocks
par(mar = c(5, 10, 6, 2)) # Set big margins
image(z = infected.arranged, y = 1:6,

      # Variety blocks are 1 unit wide, centered at 0.5, 1.5, etc
      x = seq(0.5, 5.5, 1/6),

      # Use black for 0, and use heatmap colors (red-orange-yellow-white) for 1 to 30
      col = c('black', heat.colors(30)), zlim = c(0, 30),
      xlab = '', ylab = '', yaxt = 'n', xaxt = 'n') # Don't automatically create axes or labels

# Use white line segments to visually separate the varieties
segments(x0 = 1.5:4.5, y0 = 0.5, y1 = 6.5, col = 'white')

# Place a title at the top, and label Varieties, treatment:status levels, and blocks around the image
title('Infection Counts', line = 4)
axis(3, labels = rep(levels(fert.ordered$block), 5), cex.axis = 0.75,

     # Each block is plotted in a column with width 1/6, so put the labels in the middle
     at = seq(7/12, 5 + 5/12, 1/6))
axis(2, labels = levels(with(fert.ordered, interaction(n.trt, status))), at = 1:6, las = 2)
axis(1, labels = levels(fert.ordered$variety), at = 1:5)

# Legend
par(mar = c(5, 1, 6, 2))
image(y = seq(-0.5, 30.5, 1), z = matrix(0:30, nrow = 1), axes = FALSE, ylab = '',
      col = c('#000000', heat.colors(30)), zlim = c(0, 30))
title('Legend', line = 1.5)
axis(4)
@
\end{figure}

\noindent If information about the physical layout of the field is
available, it would be possible to construct a similar heatmap to look
for spatial patterns. If correlation between nearby plots is seen,
this should be included in the model.

\noindent Another option for examining the data graphically is the
\verb|itableplot| function in the \verb|tabplot| package. This is a
useful interactive tool for visualizing large datasets and identifying
patterns.

\begin{figure}
\caption{Example \texttt{itableplot} output.}
<<tabplot,echo=FALSE,warning=FALSE,message=FALSE>>=
tableplot(fert.ordered, select = c(infected, variety, block, n.trt, status),
          nBins = 30, sortCol = variety)
@
\end{figure}

\subsection{Model Fitting and Possible Issues}
\noindent The
<<models1,warning=FALSE,results='hide',cache=TRUE>>=
glmm3way.sim <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                      control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 20000)),
                      family = binomial, data = fert.sim1)
@

<<models2,warning=FALSE,results='hide',cache=TRUE>>=
noVar1.sim <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                      control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 20000)),
                      family = binomial, data = fert.sim1, subset = variety!='Var1')
@

<<models3,warning=FALSE,results='hide',cache=TRUE>>=
glmm3way.sim <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                      control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 20000)),
                      family = binomial, data = fert.sim1)
@

<<models4,echo=FALSE,warning=FALSE,results='hide',cache=TRUE>>=
glmm3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                  control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert2)
#summary(glmm3way)

glmm2way <- glmer(cbind(infected, total) ~ variety*n.trt + variety*status2 +
                    n.trt*status2 + (1|block/row),
                    control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert2)
#summary(glmm2way)

#Anova(glmm3way)
#Anova(glmm2way)
#anova(glmm2way, glmm3way)
@

<<plots1,eval=FALSE,echo=FALSE,warning=FALSE,message=FALSE,out.height='0.5\\linewidth',fig.height=5>>=
par(mfrow = c(1, 2))
plot(resid(glmm3way) ~ predict(glmm3way), col = variety, data = fert.sim1)
plot(resid(glmm3way) ~ fitted(glmm3way), col = variety, data = fert.sim1)
legend('topright', pch = 1, col = 1:5,
       legend = levels(fert.sim1$variety))
@

<<nosnms,cache=TRUE>>=
table(fert2$infected==0,fert2$variety)

noSNMS3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                    (1|block/row), subset = variety != 'SNMS',
                    control = glmerControl(optimizer = 'bobyqa'),
                    family = binomial, data = fert2)
summary(noSNMS3way)
Anova(noSNMS3way)

noSNMS2way <- glmer(cbind(infected, total) ~ variety*n.trt + variety*status2 +
                    n.trt*status2 + (1|block/row), subset = variety != 'SNMS',
                    control = glmerControl(optimizer = 'bobyqa'),
                    family = binomial, data = fert2)
summary(noSNMS2way)
Anova(noSNMS2way)

anova(noSNMS2way, noSNMS3way)
@

<<plots3,eval=FALSE,echo=FALSE,warning=FALSE,message=FALSE,out.height='0.5\\linewidth',fig.height=5>>=
par(mfrow = c(1, 2))
plot(resid(noSNMS3way) ~ predict(noSNMS3way), col = variety,
     data = fert2, subset = variety!='SNMS')
plot(resid(noSNMS3way) ~ fitted(noSNMS3way), col = variety,
     data = fert2, subset = variety!='SNMS')
legend('topright', pch = 1, col = 1:5,
       legend = levels(fert2$variety))
@

\begin{figure}
\caption{The fitted model can be nicely summarized with plots from the \texttt{effects} package.}
<<plots2,echo=FALSE,warning=FALSE,message=FALSE,out.height='0.5\\linewidth',fig.height=5>>=
require(effects)
plot(allEffects(glmm3way.sim), multiline = TRUE, type = 'response', se = TRUE,
     ci.style='bars', x.var = 'n.trt', rotx = 45, layout = c(5, 1),
     main = 'Variety:Treatment:Inoculation Effects Plot',
     xlab = 'Nitrogen Application Date', ylab = 'Infection Probability',
     key.args = list(title = 'Inoculation Status'))
@
\end{figure}

\subsection{Model Refinement}

\subsection{Interpretation}

\section{Scope of inference}
\noindent

\section{Additional Comments}
\noindent While very low infection rates are, in practice, a desirable
result, they present technical and computational problems for analysis.
In similar future studies, random sampling should be used within each
plot and the sample size should be large enough that the researchers
expect at least one infected leaf to be found in each plot.

\section{Appendix: R Code}
\noindent Since Nar expressed a desire for assistance with R, we have
included the code we used to create the output for this report. Note
that we used a simulated dataset.

\subsection{Data Preprocessing}
\noindent We created variable with a unique level for each row, since row
it was not included in the dataset. We also changed the baseline levels
of \verb|n.application| and \verb|status|.
<<fakepre,eval=FALSE>>=
@

\subsection{\texttt{qplot}}
<<interactionplot,eval=FALSE>>=
@

\subsection{Heatmap}
Nar only needs to change \verb|fert.sim1| inside the \verb|arrange| call
to the name of his data frame. The rest of this code is self-contained and
will create the heatmap with appropriate labels for varieties, treatments,
and blocks.
<<heatmap,eval=FALSE>>=
@

\subsection{Summary Effects Plot}
<<plots2,eval=FALSE>>=
@

\section{References}
%\noindent Ramsey, F.L., Schafer, D.W. (2013). {\it The Statistical Sleuth: A Course in Methods of Data Analysis, Third Edition}. Boston, MA: Brooks/Cole, Cengage Learning.\\

\noindent Gelman, A. and Hill, J. (2007). {\it Data Analysis Using
Regression and Multilevel/Hierarchical Models}. New York, NY: Cambridge
University Press.\\

%\noindent Philippi, Tom. "Topic 2(b) GRTS Spatial Sampling (for Monitoring)." GRTS Spatial Sampling (for Monitoring). Web. 17 Feb. 2015. \\
%http://science.nature.nps.gov/im/datamgmt/statistics/r/advanced/grts.cfm. \\

%\noindent Barbour, Christopher, and Wright, Wilson. ``Consulting Report''. Manuscript, 2014.

%\noindent Ramsey, F.L., Schafer, D.W. (2013). {\it The Statistical Sleuth: A Course in Methods of Data Analysis, Third Edition}. Boston, MA: Brooks/Cole, Cengage Learning.\\

\noindent Sloderbeck, J., Michaud, P., Whitworth, Robert. "Wheat Pests." CurlMite. Kansas State University, 1 May 2008. Web. 18 Sept. 2015. \verb|http://entomology.k-state.edu/extension/insect-information/crop-pests/wheat/curlmite.html|.

\end{document}